---
title: "SIG Project Real-Time Imputation"
author: "Steven Nijman, Thomas Debray, Maarten van Smeden, Gerko Vink, Hanne Oberman"
---

# Aim

This document contain the set-up of our SIG project titled "An evaluation of 'real-time' missing data handling in machine learning and prevailing statistical models". The aim is to compare different strategies for developing prediction models that can handle the presence of missing values real time in a single patient.

# Data Generating Mechanism

```{r setup, message = FALSE, warning = FALSE, include = FALSE}
# environment
library("dplyr") #for the pipe operator
library("ggcorrplot") #for easy visualization of correlations
library("mvtnorm") #for multivariate normal distributions
library("pROC") #for easy calculation of the auc
library("mice") #for missing data stuffs

# data
load("./Data/varcov.RData") #variance-covariance matrix of the SMART dataset

# functions
source("./R/DGM.R") #data generating mechanism
```

We will use a prediction model with: 

- 1 binary outcome ($Y$),
- 10 continuous predictors ($X_1$, $X_2$, ..., $X_{10}$).

### Q: Do we need categorical predictors as well? 

--> A: No, not necessary. But if it's not too complex, add them later. Leave them for now. 

Our sample will include at least these 11 variables, with a sample size of 10.000 cases. 

### Q: Do we want extra 'support' variables ($Z$) in our sample that are not included in the prediction model?

--> A: Not yet. For now we want the ML and regression models to have the same info.

Initially, we were going to use the SMART data as the basis for our predictor space. But this would result in the same limitations as described in Nijman et al. (2021; i.e., low correlations between predictor variables). 

These correlations between variables from the SMART dataset are visualized below. 


```{r SMART}
# we'll use the SMART data to define the relations between predictors
# the variance-covariance matrix is stored in the object varcov
ggcorrplot::ggcorrplot(cov2cor(varcov), hc.order = TRUE, type = "lower", lab = TRUE)
```

So instead, we'll create our own. --> Note: maybe add more variance-covariance matrices for sensitivity analyses


```{r varcov}
# create a variance-covariance matrix with p predictors
set.seed(11)
p <- 10
sigma <- diag(p)
sigma[lower.tri(sigma)] <- sigma[upper.tri(sigma)] <- runif(p*(p-1)/2, -.9, .9) #now too extreme, rather make them 0.2-0.4 or have different options
ggcorrplot::ggcorrplot(cov2cor(sigma), hc.order = TRUE, type = "lower", lab = TRUE)
```


```{r}
# let's generate some data (note that this part is still based on the SMART data predictor space!)
n <- 10000
dat <- generate_sample(n, varcov)
glimpse(dat)

# check prevalence and auc
out <- check_characteristics(dat)
```

The prevalence of the outcome is `r out$prevalence`. The C index/AUC of the model is `r out$auc`. Note: make the C index 0.75.

### Q: What other characteristics do we need to check to see if the data suits our purpose?

--> A: With tree based methods, we'll need more interesting interactions

We added interaction terms between the first predictor and all others to create a need for ML models.

### Q: Is 10 interactions enough to make the data more realistic/complex?

--> A: Yes, make them a factor 10 or even 20 smaller than main effects. Add a spline effect as well.

### Q: Are additional non-linear effects needed (e.g., log-transformed predictors)?

--> A: Yes.

```{r}
# now add some interaction terms (actually 1 squared term, 9 interactions)
dat <- generate_sample(n, varcov, interaction = TRUE)
# check prevalence and auc
out <- check_characteristics(dat)
```

The prevalence of the outcome now is `r out$prevalence`. The C index/AUC of the model is `r out$auc`.

The next step is to ampute the complete set using several missing data patterns. Note that this part of the simulation pipeline is still very much under construction. 

### Q: Should the missing data pattern be equal for all three missing data mechanisms?

--> A: Yes. Some variables are missing more often than others. For example, lab measures are missing more often. We'll create 3 or 4 patterns where joint missingness occurs in 3 variables. 50-75% of the rows has at least one of the sets of 3 variables missing.

### Q: Should the outcome variable contain missingness? It won't be observed anyways, right?

--> A: Indeed, it will be predicted.

### Q: What is meant with MAR situation 1 and 2 in the Word doc from our last meeting?

--> A: Not relevant anymore.

```{r}
# create MCAR missingness
MCAR <- dat %>% 
  mice::ampute(mech = "MCAR")
MCARmd <- mice::md.pattern(MCAR$amp)

# create MAR missingness
MAR <- dat %>% 
  mice::ampute(mech = "MAR")
MARmd <- mice::md.pattern(MAR$amp)

# create MNAR missingness for the strongest predictor
dat %>% 
  glm(Y ~ ., family = "binomial", data = .) %>% 
  broom::tidy() %>% 
  dplyr::arrange(desc(abs(estimate)))
# the strongest predictor currently is X4
MNAR <- dat %>% 
  mice::ampute(mech = "MNAR")
MNARmd <- mice::md.pattern(MNAR$amp)
```


# Estimands

<!-- discr (c of auc) en calibr: interc en slope -->
<!-- mse van de predicties -->
<!-- verschil tussen voorspelde uitkomst en ware uitkomst (vanwege simulatie hier nu mogelijk) -->

<!-- sensitiviteit en specificiteit niet belangrijk hier -->


# Planning

- Steven and Hanne meet every week

- The entire team meets every other week, at a fixed moment in the week (e.g., wednesday afternoon)

