---
title: "SIG Results"
author: "Hanne Oberman"
date: "7-10-2021"
output:
  html_document:
    toc: yes
    theme: paper
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE)

# packages
library(dplyr)
library(ggplot2)
# library(patchwork)

# figure labels and colors
meth_ord <- c("CMI+FLR", "SDI+FLR", "MDI+FLR", "BOS+FLR", "CMI+RF", "SDI+RF", "MDI+RF", "BOS+RF", "SS+RF")
meth_lab <- c("JMI-CM", "JMI-SD", "JMI-MD", "PS", "JMI-CM", "JMI-SD", "JMI-MD", "PS", "SS")
meth_col <- c("#1269b0", "#81c454") %>% setNames(c("FLR", "RF"))
miss_lab <- c("JMI-CM", "JMI-SD", "JMI-MD", "PS", "SS")
plot_lab <- c(miss_lab[-5], miss_lab)

```

# Reference Performance

On the complete development set data, we observe the following performance of flexible logistic regression and random forest prediction models in terms of calibration and discrimination. The grey lines in Figure XYZ represent perfect calibration.

```{r reference}
# load data
ref_perf <- readRDS("../Results/reference_performance.RDS")

# table
mean_ref <- ref_perf %>% 
  group_by(method) %>% 
  summarise(auc = mean(auc), int = mean(intercept), slo = mean(slope), rmse = mean(rmse), brier = mean(brier), mae = mean(mae))

# plot
ref_long <- ref_perf %>% tidyr::pivot_longer(cols = everything()[-1]) %>% 
  mutate(name = factor(name, levels = c("intercept", "slope", "auc", "rmse", "brier", "mae"), labels = c( "Intercept", "Slope", "AUC", "RMSE", "Brier", "MAE"), ordered = TRUE)) 
# ref_long$vline <- NA
# ref_long[ref_long$name == "Intercept", "vline"] <- 0 
# ref_long[ref_long$name == "Slope", "vline"] <- 1

ref_long %>%  ggplot(aes(x = value, y = method, fill = method)) +
  #geom_vline(aes(xintercept = vline), color = "grey", linewidth = 2) +
  geom_boxplot() +
  facet_wrap(~name, scales = "free") +
  scale_y_discrete(limits = c("RF", "FLR")) + 
  scale_fill_manual(values = meth_col) +
  theme_classic() +
  labs(x = "", y = "")
```

On complete data, the flexible logistics regression and random forest prediction models have equivalent calibration:

- FLR intercept closer to 0 than RF (`r mean_ref$int %>% round(.,3)`, respectively);

- FLR slope further from 1 than RF (`r mean_ref$slo %>% round(.,3)`, respectively).

The flexible logistic regression prediction model has somewhat better discrimination than the random forest:

- FLR C-index higher than RF (`r mean_ref$auc %>% round(.,3)`, respectively).


# Simulation Results

```{r results}
# load data
performance <- readRDS("../Results/performance.RDS")

# plot function
plot_perf <- function(all_perf, metric){
  all_perf[, c("Method", "Model", "Strategy", metric)] %>% 
  setNames(c("Method", "Model", "Strategy", "metric")) %>% 
  ggplot(aes(x = metric, y = Method, fill = Model)) +
  geom_boxplot() +
  scale_y_discrete(limits = rev(meth_ord), labels = rev(plot_lab)) +
  scale_fill_manual(values = meth_col) +
  theme_classic() +
  labs(x = metric, y = "Missing data strategy")
}
```


## Calibration

```{r calibration}
plot_perf(performance, "Intercept") + list(geom_vline(xintercept = 0, color = "grey", linewidth = 2))
plot_perf(performance, "Slope") + list(geom_vline(xintercept = 1, color = "grey", linewidth = 2))
```

Aim: Intercept should be zero, slope should be one.

Summary:

- Best performance in terms of intercept: BOS+FLR, MDI+RF, MDI+FLR

- Worst performance in terms of intercept: SDI+FLR, SS+RF (wide range), SDI+RF, BOS+RF (CMI+RF = mixed bag)

- Best performance in terms of slope: BOS+FLR, CMI+FLR, MDI+RF and MDI+FLR

- Terrible performance in terms of slope: SDI+FLR, SDI+RF, BOS+RF


## AUC

```{r auc, message=FALSE, warning=FALSE}
plot_perf(performance, "AUC")
```

Aim: As close to one as possible. Method with the highest AUC has the best discrimination between patients with and without the outcome (Y).

Summary:

- Best performance in terms of AUC: SDI+RF (???), MDI+RF/CMI+FLR, MDI+FLR 

- Worst performance in terms of AUC: SDI+FLR, CMI+RF, SDI+RF, BOS+RF 


## MAE

```{r mae, message=FALSE, warning=FALSE}
plot_perf(performance, "MAE")
```

Aim: As close to zero as possible. Method with the lowest MAE has the best approximation of the true probability of Y.

Summary:

- Best performance in terms of MAE: SS+RF(???), BOS+FLR, CMI+FLR

- Worst performance in terms of MAE: CMI+RF, SDI+FLR,, BOS+RF


## RMSE

```{r RMSE}
plot_perf(performance, "RMSE")
```

Aim: As close to zero as possible. Method with the lowest RMSE has best recovered the original probability of Y.

Summary:

- Best performance in terms of RMSE: SDI+RF (???), MDI+BOS, MDI+FLR, CMI+FLR (BOS+FLR = mixed bag)

- Worst performance in terms of RMSE: SDI+FLR, CMI+RF, SS+RF (very wide range) and BOS+RF 


## Brier score

```{r brier}
plot_perf(performance, "Brier")
```

Aim: As close to zero as possible. Method with the lowest Brier score has best recovered the original observed Y.

Summary:

- Best performance in terms of Brier score: SDI+RF, MDI+RF, MDI+FLR, CMI+FLR

- Worst performance in terms of Brier score: SDI+FLR, CMI+RF, SS+RF (wide range) and BOS+RF 





## Full Figure

```{r all, fig.width=10}
# load data
performance <- readRDS("../Results/performance.RDS")

# make wide format long
perf_long <- performance %>% 
  tidyr::pivot_longer(cols = everything()[-c(1, 8, 9)], names_to = "Metric", values_to = "Performance") %>% 
  mutate(
    Metric = factor(Metric, levels = c( "RMSE", "Brier", "MAE", "AUC", "Intercept", "Slope"), labels = c("RMSE", "Brier score", "MAE",  "C-statistic",  "CITL", "Calibration slope"), ordered = TRUE),
    Strategy = factor(Strategy, levels = c("CMI", "SDI", "MDI", "BOS", "SS"), labels = miss_lab, ordered = TRUE)
    )
perf_long$vline <- NA
perf_long[perf_long$Metric == "CITL", "vline"] <- 0 
perf_long[perf_long$Metric == "Calibration slope", "vline"] <- 1

# plot
perf_long %>% filter(Metric != "MAE") %>% 
ggplot(aes(x = Performance, y = Strategy, fill = Model)) +
  geom_vline(aes(xintercept = vline), color = "grey") +
  geom_boxplot() +
  scale_y_discrete(limits = rev(miss_lab)) +
  facet_wrap(Metric~Model, scales = "free", ncol = 2, strip.position = "bottom",  dir = "h", labeller = labeller(Model = c("FLR" = " ", "RF" = " "))) +
  scale_fill_manual(values = meth_col, name = "Prediction model") +
  theme_classic() +
  theme(strip.background = element_blank(), strip.placement = "outside", legend.position = "bottom") +
  labs(x = "", y = "")

```

## Summary Table

```{r average}
perf_table <- performance %>% 
  select(-Method) %>% 
  group_by(Model, Strategy) %>% 
  summarise(across(everything(), mean),
            across(everything(), round, 3))

perf_table %>%
   write.table(file = "../Results/perf.csv", sep = ";", row.names = FALSE)

knitr::kable(perf_table,
                  format = "html",
                  booktabs = TRUE,
                  escape = FALSE,
                  align = c("l", "l", rep("c", 5)),
)
```

## Calibration plots

```{r cal-plot}
# load data
pred_dat <- readRDS("../Results/results.RDS")[[1]][1:1000,]
pred_long <- tidyr::pivot_longer(pred_dat, cols = everything()[-c(1:2)]) %>% 
  mutate(
    model = case_when(stringr::str_detect(name, "FLR")~"FLR", TRUE~"RF"),
    miss = stringr::str_remove(name, "[^+]*$"),
    miss = stringr::str_remove(miss, "[+].*"),
    miss = factor(miss, levels = c("CMI", "SDI", "MDI", "BOS", "SS"), labels = miss_lab, ordered = TRUE)) 

plot_dat <- pred_long %>%
  group_by(name) %>%
  mutate(bin = ntile(value, 10)) %>%
  group_by(name, bin) %>%
  mutate(
    n = n(),
    bin_pred = mean(value),
    bin_prob = mean(Y_prob),
    se = sqrt((bin_prob * (1 - bin_prob)) / n),
    ul = bin_prob + 1.96 * se,
    ll = bin_prob - 1.96 * se
  ) %>%
  ungroup() 

plot_dat %>%
  filter(name != "SDI+FLR" & name != "SDI+RF") %>% 
    ggplot() +
    geom_density(aes(x = value, y=..scaled..), color = "white", fill = "grey", alpha = 0.25) +
    geom_abline(linetype = "dashed", alpha = 0.5) +
    geom_smooth(aes(x = value, y = Y_prob, color = model), se = FALSE, method = "loess") +
    geom_pointrange(aes(x = bin_pred, y = bin_prob, ymin = ll, ymax = ul), size = 0.25, color = "black") +
    scale_y_continuous(limits = c(0, 1), breaks = seq(0, 1, by = 0.1)) + #, sec.axis = dup_axis(name = "Density")) +
    scale_x_continuous(limits = c(0, 0.6), breaks = seq(0, 1, by = 0.1)) +
    scale_color_manual(values = meth_col, name = "Prediction model") +
    facet_grid(miss~model) +
    labs(
      x = "Predicted probability", 
      y = "Actual probability", 
      #title = meth_lab[which(meth_ord == meth)]
      ) +
    theme_classic() +
  theme(legend.position = "bottom")


# # loosely based on https://darrendahly.github.io/post/homr/
# # for testrun: meth <- "CMI+FLR"
# plot_cali <- function(pred_dat, meth, ...){
#   plot_dat <- mutate(pred_dat, bin = ntile(get(meth), 10)) %>% 
#           # Bin prediction into 10ths
#       group_by(bin) %>%
#       mutate(n = n(), # Get ests and CIs
#              bin_pred = mean(get(meth)), 
#              bin_prob = mean(Y_prob), 
#              se = sqrt((bin_prob * (1 - bin_prob)) / n), 
#              ul = bin_prob + 1.96 * se, 
#              ll = bin_prob - 1.96 * se) %>%
#       ungroup() 
# 
#   gg_cali <- plot_dat %>%
#     ggplot() +
#     geom_density(aes(x = x, y=..scaled..), color = "white", fill = "grey", alpha = 0.25, data = pred_dat %>% select(x = meth)) +
#     geom_abline(linetype = "dashed", alpha = 0.5) +
#     geom_smooth(aes(x = get(meth), y = Y_prob),
#                     color = ifelse(grepl("FLR", meth), "#1269b0", "#81c454"), se = FALSE, method = "loess") +
#     geom_pointrange(aes(x = bin_pred, y = bin_prob, ymin = ll, ymax = ul), size = 0.25, color = "black") +
#     scale_y_continuous(limits = c(0, 1), breaks = seq(0, 1, by = 0.1)) + #, sec.axis = dup_axis(name = "Density")) +
#     scale_x_continuous(limits = c(0, 0.6), breaks = seq(0, 1, by = 0.1)) +
#     scale_color_manual(values = meth_col) +
#     labs(
#       x = "Predicted probability", 
#       y = "Actual probability", 
#       title = meth_lab[which(meth_ord == meth)]) +
#     theme_classic()
# }   
# pred_plots <- purrr::map(meth_ord, ~plot_cali(pred_dat, .x))
# 
# # TODO: make one fig?
# # TODO: add histogram of pred prob?
# 
# # JMI-CM en SD naar supplement
# # de rest in één figuur
# 
# library(patchwork)
# 
# (pred_plots[[1]] + pred_plots[[5]]) / (pred_plots[[3]] + pred_plots[[7]]) / (pred_plots[[4]] + pred_plots[[8]]) / (plot_spacer() + pred_plots[[9]])

```

