---
title: "Supplementary Materials"
author: "Hanne Oberman"
date: "15-11-2021"
output: word_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
library(dplyr)
library(ggplot2)
```

# Methods


## Variance-covariance matrix

The variance-covariance matrix of the predictor space    

$$
{\Sigma} = \left[ \begin{array}{cccccccccc} 
1.05  & -0.12 & 0.04  & -0.29 & 0.29  & -0.17 & 0.01  & 0     & -0.01 & -0.07 \\
-0.12 & 1.08  & -0.31 & 0.26  & 0.08  & -0.03 & -0.04 & -0.11 & -0.17 & 0.3   \\
0.04  & -0.31 & 1.08  & -0.19 & 0.01  & -0.29 & 0.2   & 0.07  & -0.18 & -0.15 \\
-0.29 & 0.26  & -0.19 & 1.07  & -0.2  & 0     & -0.12 & 0.01  & -0.19 & -0.04 \\
0.29  & 0.08  & 0.01  & -0.2  & 1.08  & -0.25 & -0.14 & 0.02  & 0.15  & -0.32 \\
-0.17 & -0.03 & -0.29 & 0     & -0.25 & 1.08  & -0.13 & -0.04 & -0.29 & 0.01  \\
0.01  & -0.04 & 0.2   & -0.12 & -0.14 & -0.13 & 1.04  & -0.16 & -0.17 & 0.18  \\
0     & -0.11 & 0.07  & 0.01  & 0.02  & -0.04 & -0.16 & 1.02  & 0.1   & -0.19 \\
-0.01 & -0.17 & -0.18 & -0.19 & 0.15  & -0.29 & -0.17 & 0.1   & 1.08  & 0.15  \\
-0.07 & 0.3   & -0.15 & -0.04 & -0.32 & 0.01  & 0.18  & -0.19 & 0.15  & 1.08
\end{array} \right].
$$


## Regression coefficients

Regression coefficients of the main effects of the predictors and interaction effects

$$
\begin{array}{lllccccccccc}
\beta     &= &[&-0.27	&0.53	&-0.97	&-0.05	&0.62	&-0.52	&0.53	&-0.61	&0.17	&-0.55&]\\
\beta^*    &= &[&0.06	&0.04	&-0.02	&-0.02	&-0.06	&-0.05	&0.04	&0.05	&0.01	&-0.07&].
\end{array}
$$


## Model tuning

- FLR: glm() with natural spline with 3 degrees of freedom.

- RF: ranger::ranger() with defaults (500 trees and 3 predictors considered for each split), party::cforest() with defaults (500 trees, 5 predictors considered for each split, and 3 surrogate variables considered for each split with missingness).


## Reference performance

- FLR:

- RF:


# Results

## Performance under MNAR missingness

```{r}
# load data
performance <- readRDS("../Results/performance_MNAR.RDS")
miss_lab <- c("CMI", "SDI", "MDI", "BOS", "SS")
plot_lab <- c(miss_lab[-5], miss_lab)
meth_ord <- c("CMI+FLR", "SDI+FLR", "MDI+FLR", "BOS+FLR", "CMI+RF", "SDI+RF", "MDI+RF", "BOS+RF", "SS+RF")
meth_col <- c("#1269b0", "#81c454") %>% setNames(c("FLR", "RF"))

# make wide format long
perf_long <- performance %>% 
  tidyr::pivot_longer(cols = everything()[-c(1, 8, 9)], names_to = "Metric", values_to = "Performance") %>% 
  mutate(
    Metric = factor(Metric, levels = c("Intercept", "Slope", "AUC", "MAE", "RMSE", "Brier"), labels = c("CITL", "Calibration slope", "C-statistic", "MAE", "RMSE", "Brier score"), ordered = TRUE)
    )

# plot
ggplot(perf_long, aes(y = Method, x = Performance, fill = Model)) +
  geom_boxplot() +
  scale_y_discrete(limits = rev(meth_ord), labels = rev(c(miss_lab[-5], miss_lab))) +
  facet_wrap(~Metric, scales = "free_x", nrow = 2) +
  scale_fill_manual(values = meth_col) +
  theme_classic()

```

